From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-30 Pi+/paper/ufin.tex
Write a segue into section 4

In the previous section, we have picked a particular encoding of finite types: as sequences of elements in some
canonical order.

This is not a ``free'' presentation. Some properties of the encoding might pollute our semantics. We want to make that
any theorems / properties of the semantics we prove apply to the source language (under various interpretations).

This motivates us to use a different semantic representation: a univalent universe of finite types which contains the
finite types and nothing but the finite types and is defined such as alternative representations of finite types are
identified. This guarantees that we can get a proper, sound and complete, semantics for our language.

>>>>

In the previous section, we gave a sound denotational semantics for $\PiLang$ using the groupoid of finite sets and
bijections $\BFin$. Types are interpreted as objects in $\BFin$, and 1-combinators are interpreted as isomorphisms in
$\BFin$. We also showed that if there is a 2-combinator between two 1-combinators, then their interpretations are
extensionally equal in $\BFin$.

However, we want to go a step further, and claim completeness for this semantics, that is, we want to quote back from
the semantics to the syntax. Given a morphism in $\BFin$, we want to produce a 1-combinator in $\PiLang$, and given two
equal morphisms in $\BFin$, we want to produce a 2-combinator witnessing the equality of the quoted 1-combinators.

The groupoid $\BFin$ that we presented, is strict, that is, two equal permutations are extensionally identified, without
an explicit witness for the equality. Our first step is to weaken this groupoid $\BFin$ into a weak 2-category. In a
weak category, equational axioms of the category are replaced by higher invertible cells. For example, there are higher
invertible cells for the left and right units of composition and associativity. The higher cells have their own
coherences as well, the 2-cells can be composed with 1-cells by whiskering, and so on.

HoTT provides a rich internal language for working with weak higher groupoids, using the ``Types are weak
$\infty$-groupoids'' correspondence. We will make this precise by observing that a 1-groupoid in HoTT gives a locally
strict $(2,0)$-category, using the identity type for the higher cells. The groupoid $\BFin$ can be written in HoTT as a
type $\UFin$, which is a 1-groupoid.

Secondly, we would like to like to give a syntactic presentation for this groupoid, by describing exactly how the
1-cells and 2-cells are generated. Note that, in $\BFin$, the set of bijections on a finite set of cardinality $n$ has a
group structure, the identity function is the neutral element, group multiplication is given by function composition,
and the group inverse is given by inverting the bijection. This is the automorphism group of the finite set, which is
also known to be the finite symmetric group $\Sn[n]$.

Groups can be defined by listing the carrier set and the operations of the group, or they can be described using a group
presentation, which gives the generators and relations of a group. The last step of our technical development is to use
a group presentation for $\Sn$ to give a presentation for the higher cells of the groupoid $\UFin$. Using this
characterisation, we will obtain our completeness result.

% In the previous section, we have picked a particular encoding of finite types: as sequences of elements in some
% canonical order.

% This is not a ``free'' presentation. Some properties of the encoding might pollute our semantics. We want to make that
% any theorems / properties of the semantics we prove apply to the source language (under various interpretations).

% This motivates us to use a different semantic representation: a univalent universe of finite types which contains the
% finite types and nothing but the finite types and is defined such as alternative representations of finite types are
% identified. This guarantees that we can get a proper, sound and complete, semantics for our language.
